Technical Report

### 2018-03-20 fully-abstract-comp/technical-report.tex
Comment out unneeded case in value relation for sealed caps.
Remove obsolete title.

\subsection{Future world}
\subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
% {\color{DarkGreen}
The public part of the future world relation stays the same:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \pubft W.i(r) }
            { W' \future W }
\end{mathpar}
The private future world relation becomes:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
            { W' \future W }
\end{mathpar}
where we take ``is non-empty'' to mean the following:

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
The argument for having one relation relate pairs of pairs of capabilties and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
\begin{align*}
  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm \\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}\\
  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}
\end{align*}

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    \; \middle| \;
    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l}
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l }
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps. It expresses that any reasonable executable capability is safe to use as the pc capability. 
\dominique{I'm following the example of the previous FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
\begin{theorem}[FTLR]
  \label{thm:ftlr}
  \[
    \begin{array}{l}
      \forall n,W,\lin,\baddr,\eaddr,\aaddr \ldotp \\
      \quad \left(
      \begin{array}{l}
        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge \\
        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \wedge \\
        \text{the compilation assumptions on $n$, $W$, ...}
      \end{array}
      \right)\\
      \quad\Rightarrow 
      \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
                      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)

    \end{array}
  \]
  \dominique{6-3-2018: I added $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
  \begin{definition}[The compilation assumptions]
    No call uses $\rtmp{1}$ for jumping.\\
    ...
  \end{definition}

>>>>

\subsection{Future world}
% {\color{DarkGreen}
The public part of the future world relation stays the same:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \pubft W.i(r) }
            { W' \future W }
\end{mathpar}
The private future world relation becomes:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
            { W' \future W }
\end{mathpar}
where we take ``is non-empty'' to mean the following:

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
The argument for having one relation relate pairs of pairs of capabilties and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
\begin{align*}
  \lrexjg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm \\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}\\
  \lreg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}
\end{align*}

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o)
      % 20-3-2018: I expect the next case (commented out below) is never used and hence unneeded.
      % \wedge \\
      % \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    \; \middle| \;
    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l}
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
      \begin{array}{r l l }
        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps. It expresses that any reasonable executable capability is safe to use as the pc capability. 
\begin{theorem}[FTLR]
  \label{thm:ftlr}
  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
  If
  \begin{itemize}
  \item $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
  \end{itemize}
  and one of the following sets of requirements holds:
  \begin{itemize}
  \item \begin{itemize}
    \item $[\baddr,\eaddr] \subseteq \ta$
    \item the compilation assumptions on $n$, $W$, ...
    \item the full abstraction conditions (difference with the previous point?)
    \end{itemize}
  \item
    \begin{itemize}
    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
    \end{itemize}
  \end{itemize}
  Then
  \[
    \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)
  \]
  \dominique{6-3-2018: I added $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
  \begin{definition}[The compilation assumptions]
    No call uses $\rtmp{1}$ for jumping.\\
    ...
  \end{definition}
